```{r}
library(tidyverse)
library(nflfastR)
```

### Terry McLaurin targets by half

### Load play-by-play data

```{r}
pbp_2022 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2022.rds")
```

### 2022: First half plays, WAS possession, pass play, McLaurin target, first three games

```{r}
mclaurin_2022 <- pbp_2022 |>
  filter(
    posteam == 'WAS',
    game_half == 'Half1',
    play_type == 'pass',
    receiver_player_name == 'T.McLaurin',
    week <= 3
  )
```

### 2023

```{r}
pbp_2023 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2023.rds")

mclaurin_2023 <- pbp_2023 |>
  filter(
    posteam == 'WAS',
    game_half == 'Half1',
    play_type == 'pass',
    receiver_player_name == 'T.McLaurin',
    week <= 3
  )
```

### 2024

```{r}
pbp_2024 <- read_rds("https://github.com/nflverse/nflverse-data/releases/download/pbp/play_by_play_2024.rds")

mclaurin_2024 <- pbp_2024 |>
  filter(
    posteam == 'WAS',
    game_half == 'Half1',
    play_type == 'pass',
    receiver_player_name == 'T.McLaurin',
    week <= 3
  )
```

In the first halves of the first three games of the season, Terry McLaurin has been targeted for 12 or less passes, with 6 in 2022, 12 in 2023 and 7 in 2024. He is being targeted only a handful of times in each first half. As a wide receiver, McLaurin should be getting the ball more often. This either speaks to him not positioning himself correctly or the team not running the best plays.
