---
title: "Women's Soccer Data"
output: html_notebook
---

We'll use this notebook to explore some NCAA women's soccer data and [our questions](https://docs.google.com/spreadsheets/d/1kTSrnl1URmcexIWLDMWqDnp3q-c2OS45uifnhjp10n4/edit#gid=0). First, let's load the tidyverse library:

```{r}
library(tidyverse)
```

We'll want to start with 2023 match data, which is located in [this GitHub repository](https://github.com/dwillis/NCAAWomensSoccerData). We can load the match data directly from GitHub using `read_csv`:

```{r}
matches_2023 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2023.csv")
```

Let's take a look at this data to see what it contains, using `head` and `glimpse`:

```{r}
head(matches_2023)
glimpse(matches_2023)
```

Now, onto the questions using aggregates, which means `group_by` and `summarize`:

1.  How many shots in total has Maryland attempted and how many have been on goal? Calculate a percentage of shots on goal. **Maryland has attempted 45 shots with 22 on goal, giving a percentage of about 49% o**

```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")

matches_2024 |>
  group_by(team) |> filter(team_id==392) |>
  summarise(
    total_shots = sum(sh_att),
    shots_on_goal = sum(so_g),
    pct_goal_shots = shots_on_goal/total_shots
  )
```

2.  What about the same shot stats from last season's first four games? You'll need to load the matches from 2022 - just change the season in the URL - and filter by date to find the first four games. **The first four games of the season were played by the teams in the table with the Akron Zips having the highest percentage of shots on goal in their game with 5 of 7 shots on target.**

```{r}
matches_2022 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2022.csv")

matches_2022 |>
  arrange(date) |> slice(1:4) |>
  group_by(team) |>
   summarise(
    total_shots = sum(sh_att),
    shots_on_goal = sum(so_g),
    pct_goal_shots = shots_on_goal/total_shots
  )

```

3.  What about draws - Maryland started 2022 with 4 ties. Have any other teams had as many this season? **Yes, several teams have had more than 4 draws with two teams already having 6.**

```{r}
matches_2024 <- read_csv("https://raw.githubusercontent.com/dwillis/NCAAWomensSoccerData/main/data/ncaa_womens_soccer_matchstats_2024.csv")

matches_2024 |>
  group_by(team, outcome = 'Draw') |>
  summarise(
    count = n()
  ) |>
  arrange(desc(count))
```

4.  How many total corner kicks has Maryland had so far and how many have its opponents have? Calculate the difference between the two totals for every team. **Maryland has had 18 corners this season while their opponents have had 23, meaning their opponents have had 5 more than them so far this season. Virginia Tech has the greatest corner difference so far of 42 with 48 total corners and their opponents only having 6.**

```{r}
matches_2024 |>
  group_by(team) |>
  summarise(
    total_corners = sum(corners),
    opp_corners = sum(defensive_corners),
    corner_diff = total_corners - opp_corners
  ) |>
  arrange(desc(corner_diff))
```
