**HEADLINE**

**By Marwa Barakat**

Lionel Messi, arguably one of the greatest footballers (and athletes) of all time, spent nearly his entire career at one club: FC Barcelona.

In his 17 seasons with the Culers, Messi won numerous individual and team awards and accolades, including 10 La Liga titles. Despite playing a couple of seasons at PSG and Inter Miami since leaving, he remains the club's top scorer in its history.

A few years ago, StatsBomb [released the entirety of Messi's La Liga career data for public use.](https://statsbomb.com/news/statsbomb-release-free-messi-data-all-seasons-from-2004-05-2020-21-now-available/) Using this data, I investigated some of Messi's most significant individual statistics, including all-time scoring and playing against rival Real Madrid – even though this could probably be a whole book.

Known for his powerful left foot, Messi has used it to score nearly all of his 18 league goals against Madrid. Despite breaking records in his 2011/12 season, scoring more than any other player in a single season from the top five European leagues, none of those league goals were scored against Madrid.

Outside of typical goal-scoring, Messi is known for scoring precise free kicks. Throughout his league career, he scored 39 free kicks – all with his special left foot.

While Messi's impact, for Barcelona and his other teams, reaches far beyond just La Liga, here is more information about this significant era in his career.

**Through StatsBomb, I loaded the necessary packages and data to create dataframes for all of Barcelona's La Liga seasons.**

```{r Packages and Data, include=FALSE}
#Load in necessary packages.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(StatsBombR))
suppressPackageStartupMessages(library(SBpitch))

#Make a dataframe for all La Liga seasons that Messi played in.
competitions <- FreeCompetitions() |>
filter(competition_id==11 & season_name!='1973/1974')

#Make a dataframe for all relevant matches from those seasons.
barca_matches <- FreeMatches(competitions) |> filter(home_team.home_team_name=='Barcelona' | away_team.away_team_name == 'Barcelona')

#Make a df for all event data from those matches.
event_data <- free_allevents(MatchesDF = barca_matches, Parallel = T) 
```

**Then, I narrowed it down to Messi's La Liga seasons and all of his actions in each game. I also added a column with each of the seasons. This showed me that the 2011/12 season had the most actions for Messi on the field.**

```{r}
#Narrow down to Messi's actions only.
messi_event <- event_data |> filter(player.id==5503) 

#This shows how many "events" he had in each season in the data. 
messi_event <- messi_event |>
  mutate(
    season_year = case_when(
      season_id == 1 ~ "2017/18",
      season_id == 2 ~ "2016/17",
      season_id == 4 ~ "2018/19",
      season_id == 21 ~ "2009/10",
      season_id == 22 ~ "2010/11",
      season_id == 23 ~ "2011/12",
      season_id == 24 ~ "2012/13",
      season_id == 25 ~ "2013/14",
      season_id == 26 ~ "2014/15",
      season_id == 27 ~ "2015/16",
      season_id == 37 ~ "2004/05",
      season_id == 38 ~ "2005/06",
      season_id == 39 ~ "2006/07",
      season_id == 40 ~ "2007/08",
      season_id == 41 ~ "2008/09",
      season_id == 42 ~ "2019/20",
      season_id == 90 ~ "2020/21"
      ))

messi_event |>
  group_by(season_year) |>
  summarise(
    total = n()
  ) |> arrange(desc(total))
```

**With the intention of looking at Messi's stats against biggest rival Madrid, I loaded a dataframe for Barca's matches against Madrid and all of the on-field events from those. I also created a df for Messi's actions only. The 2010/11 season had the most actions.**

```{r}
#Narrow down Messi's games/actions against Real Madrid.
barca_madrid <- barca_matches |>
  filter(away_team.away_team_name=='Real Madrid' | home_team.home_team_name=='Real Madrid')

clasico_event_data <- free_allevents(MatchesDF = barca_madrid, Parallel = T) 

messi_clasico <- clasico_event_data |> filter(player.id==5503)

#I did this to add the season years to the data for easier analysis.
messi_clasico <- messi_clasico |>
  mutate(
    season_year = case_when(
      season_id == 1 ~ "2017/18",
      season_id == 2 ~ "2016/17",
      season_id == 4 ~ "2018/19",
      season_id == 21 ~ "2009/10",
      season_id == 22 ~ "2010/11",
      season_id == 23 ~ "2011/12",
      season_id == 24 ~ "2012/13",
      season_id == 25 ~ "2013/14",
      season_id == 26 ~ "2014/15",
      season_id == 27 ~ "2015/16",
      season_id == 37 ~ "2004/05",
      season_id == 38 ~ "2005/06",
      season_id == 39 ~ "2006/07",
      season_id == 40 ~ "2007/08",
      season_id == 41 ~ "2008/09",
      season_id == 42 ~ "2019/20",
      season_id == 90 ~ "2020/21"
      ))

messi_clasico |>
  group_by(season_year) |>
  summarise(
    total = n()
  ) |> arrange(desc(total))
```

**After loading the necessary data, I started small. I wanted to see which seasons were Messi's most impactful by looking at shots and goals. Here we can start to see how his 2011/12 season was his best performance.**

```{r}
#This is all of his shots and goals by La Liga season
messi_sg <- messi_event |>
  group_by(season_year) |>
  summarise(
    total_shots = sum(type.name=='Shot', na.rm = TRUE),
    total_goals = sum(shot.outcome.name=="Goal", na.rm = TRUE))

#Here I did a simple bar chart to compare his season stats
messi_sg_wide <- messi_sg |>
  pivot_longer(
    cols = starts_with('total'),
    names_to = "type",
    values_to = "total"
  )

ggplot() + 
  geom_bar(
    data=messi_sg_wide, 
    aes(x=reorder(season_year, total), 
        weight= total,
        fill=type)) + 
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Lionel Messi breaks records in the 2011/12 season",
    y = "Total",
    x = "Season"
  )
```

**But in order to separate shots from goals, I made a separate chart to see which seasons actually resulted in the most goals. Again, the 2011/12 season stands out spectacularly at 50 goals, with his first season as a senior player, 2004/05, at the bottom like expected.**

```{r}
messi_season_goals <- messi_event |>
  group_by(season_year) |>
  summarise(
     total_goals = sum(shot.outcome.name=="Goal", na.rm = TRUE))

ggplot() + 
  geom_bar(
    data = messi_season_goals, 
    aes(
      x = reorder(season_year, total_goals), 
      weight = total_goals, 
      fill = ifelse(season_year == "2011/12", "highlight", "default")
    )
  ) + 
  scale_fill_manual(
    values = c("highlight" = "red", "default" = "gray")
  ) + 
  labs(
    title = "Messi breaks goal-scoring records in 2011/12 season", 
    x = "Season", 
    y = "Total Goals", 
    fill = NULL
  ) + 
  theme_minimal() + 
    theme(legend.position = "none") + 
  coord_flip()
```

**Before I could get into making graphics or visualizations on a soccer pitch, I needed to split the location columns into separate coordinates for the x and y axes.**

```{r}
#Adding these columns to all dataframes.
messi_clasico <- messi_clasico |>
   mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )

messi_event <- messi_event |>
   mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )
```

**Then, I followed the thread of Messi's performance against Madrid by looking at his non-penalty shots against the team and their expected goal value (which is generated by a StatsBomb equation). Here, we can see his left foot standing out with plenty of shots made from outside of the box.**

```{r}
shots <- messi_clasico |>
 filter(type.name=="Shot" & (shot.type.name!="Penalty" | is.na(shot.type.name)))

shotmapxgcolors <- c("#192780", "#2a5d9f", "#40a7d0", "#87cdcf", "#e7f8e6", "#f4ef95", "#FDE960", "#FCDC5F",
 "#F5B94D", "#F0983E", "#ED8A37", "#E66424", "#D54F1B", "#DC2608", "#BF0000", "#7F0000", "#5F0000")

ggplot() +
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 # add penalty spot right
 annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) +
 annotate("path", colour = "black", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 # add centre spot
 annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 annotate("path", x=107.84-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 geom_point(data = shots, aes(x = location.x, y = location.y, fill = shot.statsbomb_xg, shape = shot.body_part.name),
 size = 3, alpha = 0.8) +
 labs(title = "Lionel Messi, Shot Map", subtitle = "La Liga, All Clasico Games") + #4
 scale_fill_gradientn(colours = shotmapxgcolors, limit = c(0,0.8), oob=scales::squish, name = "Expected Goals
Value") +
   scale_shape_manual(values = c("Head" = 21, "Right Foot" = 23, "Left Foot" = 24), name ="") + #6
 guides(fill = guide_colourbar(title.position = "top"),
 shape = guide_legend(override.aes = list(size = 4, fill = "black"))) + #7
 coord_flip(xlim = c(85, 125)) +
  theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=5, hjust=0.5, vjust=0.5),
  plot.subtitle = element_text(size = 15, hjust = 0.5),
 axis.text.y=element_blank(), 
 legend.title=element_text(size=10),
 legend.text=element_text(size=10),
 legend.key.size = unit(0.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 20, colour = "black", hjust = 0.5),
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=8)) 
```

**But from here, it was important to look at his goals against Madrid. We can see the goals are spread across numerous seasons and mostly with his left foot.**

```{r}
messi_clasico_goals <- messi_clasico |>
  filter(shot.outcome.name=="Goal")

ggplot() +
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 # add penalty spot right
 annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) +
 annotate("path", colour = "black", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 # add centre spot
 annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 annotate("path", x=107.84-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 geom_point(data = messi_clasico_goals, aes(x = location.x, y = location.y, shape = shot.body_part.name, colour = season_year),
 size = 3, alpha = 0.8) +
 labs(title = "Lionel Messi's Goals vs. Madrid", subtitle = "La Liga",) + #4 +
   scale_shape_manual(values = c("Head" = 21, "Right Foot" = 23, "Left Foot" = 24), name ="") + #6 
 guides(fill = guide_colourbar(title.position = "top"),
 shape = guide_legend(override.aes = list(size = 4, fill = "black"))) + #7
 coord_flip(xlim = c(85, 125)) +
  theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=5, hjust=0.5, vjust=0.5),
  plot.subtitle = element_text(size = 15, hjust = 0.5),
 axis.text.y=element_blank(), 
 legend.title=element_text(size=10),
 legend.text=element_text(size=10),
 legend.key.size = unit(0.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 20, colour = "black", hjust = 0.5),
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=8)) 
```

**But of course, Messi is more than just a goal scorer as the player with the highest number of assists in football history. In continuing with an analysis of Clasico games, I graphed all of his assists against their rival.**

```{r}
messi_assists <- messi_clasico |>
  filter(pass.goal_assist == TRUE & player.id==5503) 

create_Pitch() +
 geom_segment(data = messi_assists, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Assists", subtitle = "La Liga,
vs. Real Madrid") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100)
```

**From this point, I was interested in making more charts like that one to display different statistics. I wanted to show all of Messi's La Liga free kick goals. However, there are simply too many for this chart to be comprehensible.**

```{r}
messi_freekick <- messi_event |>
  filter(shot.outcome.name=="Goal" & shot.type.name=='Free Kick')

messi_freekick <- messi_freekick |>
   mutate(
    shot.end_location.x = map_dbl(shot.end_location, ~ .x[1] %||% NA),
    shot.end_location.y = map_dbl(shot.end_location, ~ .x[2] %||% NA)
  )

create_Pitch() +
 geom_segment(data = messi_freekick, aes(x = location.x, y = location.y,
 xend = shot.end_location.x, yend = shot.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Freekick Goals", subtitle = "La Liga") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100) 
```

**That led me to change the chart to this one to individually show where each of the free kick goals started and what season they occurred. This created a cleaner display, showing Messi's sweet spot falling closer to the right side of the field for these goals — including an exceptional free kick from the right side of the box.**

```{r}
ggplot() +
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 # add penalty spot right
 annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) +
 annotate("path", colour = "black", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 # add centre spot
 annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 annotate("path", x=107.84-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 geom_point(data = messi_freekick, aes(x = location.x, y = location.y, shape = shot.body_part.name, colour=season_year),
 size = 3, alpha = 0.8) +
 labs(title = "Lionel Messi, Free Kick Goal Map", subtitle = "La Liga") + #4 +
   scale_shape_manual(values = c("Head" = 21, "Right Foot" = 23, "Left Foot" = 24), name ="") + #6 
 guides(fill = guide_colourbar(title.position = "top"),
 shape = guide_legend(override.aes = list(size = 4, fill = "black"))) + #7
 coord_flip(xlim = c(85, 125)) +
  theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=5, hjust=0.5, vjust=0.5),
  plot.subtitle = element_text(size = 15, hjust = 0.5),
 axis.text.y=element_blank(), 
 legend.title=element_text(size=10),
 legend.text=element_text(size=10),
 legend.key.size = unit(0.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 20, colour = "black", hjust = 0.5),
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=8))
```

**As an assister, Messi is known for long passes that push the ball forward. I wanted to create a simple chart to see long passes (which I defined as over 40 yards). This isn't digestible, but as a longtime Barca fan, I noticed plenty of these passes were toward the left side of the goal, which is where Messi's longtime teammate Jordi Alba is generally positioned.**

```{r}
messi_long_passes <- messi_clasico |>
  filter(type.name=='Pass' & pass.length>40)

create_Pitch() +
 geom_segment(data = messi_long_passes, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Long Passes", subtitle = "La Liga, All Clasico Games") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100) 
```

**So, I created a chart to look at only Messi's long passes to Alba in Clasico games, which make up for most of the left-traveling passes from the previous chart.**

```{r}
messi_alba <- messi_event |>
  filter(type.name=='Pass' & pass.recipient.name=='Jordi Alba Ramos')

messi_alba_clasico <- messi_clasico |>
    filter(type.name=='Pass' & pass.recipient.name=='Jordi Alba Ramos' & pass.length>40)

create_Pitch() +
 geom_segment(data = messi_alba_clasico, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Messi-Alba Connection", subtitle = "La Liga, Clasico") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100) 
```

**Finally, knowing Messi's range is greater and more consistent than most players, I wanted to analyze his outside of the box goals (that aren't free kicks). Here, I found that his sweet spot for these goals falls just about the center of outside the box.**

```{r}
messi_outbox <- messi_event |>
  filter(shot.type.name!='Free Kick') |>
  filter(shot.outcome.name=="Goal") |>
  filter(
    location.x<102 | location.y>62 | location.y<18
  )

ggplot() +
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 # add penalty spot right
 annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) +
 annotate("path", colour = "black", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 # add centre spot
 annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 annotate("path", x=107.84-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 geom_point(data = messi_outbox, aes(x = location.x, y = location.y, colour = season_year),
 size = 3, alpha = 0.8) +
 labs(title = "Lionel Messi, Outside the Box Goals", subtitle = "La Liga (excludes free kick goals)",) + #4 +
 guides(fill = guide_colourbar(title.position = "top"),
 shape = guide_legend(override.aes = list(size = 4, fill = "black"))) + #7
 coord_flip(xlim = c(85, 125)) +
  theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=5, hjust=0.5, vjust=0.5),
  plot.subtitle = element_text(size = 15, hjust = 0.5),
 axis.text.y=element_blank(), 
 legend.title=element_text(size=10),
 legend.text=element_text(size=10),
 legend.key.size = unit(0.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 20, colour = "black", hjust = 0.5),
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=8)) 
```

**Through this analysis (and his career trajectory), it's safe to say Messi was a one-of-a-kind La Liga player.**

**ATTEMPTED CHART TRASH**

The following are all of the visualizations or dataframes I attempted or made but felt had no place in my presentation. They took too much time and energy to fully remove from the notebook, so you can browse them below if interested. Please disregard them when grading (unless you like them).

```{r}
messi_alba_assists <- messi_event |>
    filter(type.name=='Pass' & pass.goal_assist == TRUE & pass.recipient.name=='Jordi Alba Ramos')

create_Pitch() +
 geom_segment(data = messi_alba_assists, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Messi Assisting Alba", subtitle = "La Liga") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100) 
```

```{r}
#Here I worked on a graph from the StatsBomb guide to show all of Messi's completed box passes against Madrid.
messi_passes <- messi_clasico |>
  mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )

messi_event |> allclean()

messi_passes <- messi_passes |>
  filter(type.name=="Pass" & is.na(pass.outcome.name) & player.id==5503) |>
  filter(pass.end_location.x>=102 & pass.end_location.y<=62 & pass.end_location.y >=18)

create_Pitch() +
 geom_segment(data = messi_passes, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Completed Box Passes", subtitle = "La Liga,
vs. Real Madrid") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100)
```

```{r}
#Testing out heat maps!
messi_heatmap <- messi_clasico |> 
  mutate(location.x = ifelse(location.x>120, 120, location.x),
         location.y = ifelse(location.y>80, 80, location.y),
         location.x = ifelse(location.x<0, 0, location.x),
         location.y = ifelse(location.y<0, 0, location.y))

messi_heatmap$xbin <- cut(messi_heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
messi_heatmap$ybin <- cut(messi_heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE)

messi_heatmap <- messi_heatmap |>
  filter(type.name=="Shot" | type.name=="Pressure") |>
  group_by(season_year) |>
  mutate(total_sp = n()) |>
  group_by(season_year, xbin, ybin) |>
  summarise(total_sp = max(total_sp),
    bin_sp = n(),
    bin_pct = bin_sp/total_sp,
    location.x = median(location.x),
    location.y = median(location.y)) 

library(grid)

spcolor <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51",
"#e35256", "#e76266", "#e9777b", "#ec8589", "#ec898d", "#ef9195",
"#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2",
"#f7e1e3", "#f5e2e4", "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd",
"#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98", "#798590",
"#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048",
"#11263e", "#11273e", "#0d233a", "#020c16")


ggplot(data= messi_heatmap, aes(x = location.x, y = location.y, fill = bin_pct, group =bin_pct)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = spcolor, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(0.03, -0.03)) + #6
 labs(title = "TITLE", subtitle = "La Liga, Clasico") + #7
 coord_fixed(ratio = 95/100) + 
  annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 facet_wrap(~season_year)+ #10
 guides(fill = guide_legend(reverse = TRUE)) + 
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()  #remove y axis ticks
        )
```

```{r}
#This is Messi's shots and goals in Clasico games in La Liga
messi_clasico_sg <- messi_clasico |>
  group_by(season_year) |>
  summarise(
    shots = sum(type.name=="Shot", na.rm = TRUE),
    goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) |> arrange(desc(goals))
```
