**By Marwa Barakat**

Data: StatsBomb, [https://statsbomb.com/news/statsbomb-release-free-messi-data-all-seasons-from-2004-05-2020-21-now-av](https://statsbomb.com/news/statsbomb-release-free-messi-data-all-seasons-from-2004-05-2020-21-now-available/){.uri}

```{r}
#Load in necessary packages.
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(StatsBombR))
```

```{r}
#Make a dataframe for all La Liga seasons that Messi played in.
competitions <- FreeCompetitions() |>
filter(competition_id==11 & season_name!='1973/1974')

#Make a dataframe for all relevant matches from those seasons.
barca_matches <- FreeMatches(competitions) |> filter(home_team.home_team_name=='Barcelona' | away_team.away_team_name == 'Barcelona')

#Make a df for all event data from those matches.
event_data <- free_allevents(MatchesDF = barca_matches, Parallel = T) 
```

```{r}
#Narrow down to Messi's actions only.
messi_event <- event_data |> filter(player.id==5503) 

#This shows how many "events" he had in each season in the data. 
messi_event <- messi_event |>
  mutate(
    season_year = case_when(
      season_id == 1 ~ "2017/18",
      season_id == 2 ~ "2016/17",
      season_id == 4 ~ "2018/19",
      season_id == 21 ~ "2009/10",
      season_id == 22 ~ "2010/11",
      season_id == 23 ~ "2011/12",
      season_id == 24 ~ "2012/13",
      season_id == 25 ~ "2013/14",
      season_id == 26 ~ "2014/15",
      season_id == 27 ~ "2015/16",
      season_id == 37 ~ "2004/05",
      season_id == 38 ~ "2005/06",
      season_id == 39 ~ "2006/07",
      season_id == 40 ~ "2007/08",
      season_id == 41 ~ "2008/09",
      season_id == 42 ~ "2019/20",
      season_id == 90 ~ "2020/21"
      ))

messi_event |>
  group_by(season_year) |>
  summarise(
    total = n()
  ) |> arrange(desc(total))
```

11/16 Update #1:

So far, I have narrowed down the data to make a dataframe of Messi's La Liga data from the 2004/05 season to the 2020/21 season. I also have all the La Liga data from those seasons from other teams as well. The data is very large and complex, so I just need to narrow down what I want to look at by looking through the guides they posted to see what all the columns are and what is possible. I think I'm leaning toward looking at passing or shooting in some capacity. I might even do something like just looking at Messi playing against Real Madrid to narrow the scope more.

```{r}
#Narrow down Messi's games/actions against Real Madrid.
barca_madrid <- barca_matches |>
  filter(away_team.away_team_name=='Real Madrid' | home_team.home_team_name=='Real Madrid')

clasico_event_data <- free_allevents(MatchesDF = barca_madrid, Parallel = T) 

messi_clasico <- clasico_event_data |> filter(player.id==5503)

#I did this to add the season years to the data for easier analysis.
messi_clasico <- messi_clasico |>
  mutate(
    season_year = case_when(
      season_id == 1 ~ "2017/18",
      season_id == 2 ~ "2016/17",
      season_id == 4 ~ "2018/19",
      season_id == 21 ~ "2009/10",
      season_id == 22 ~ "2010/11",
      season_id == 23 ~ "2011/12",
      season_id == 24 ~ "2012/13",
      season_id == 25 ~ "2013/14",
      season_id == 26 ~ "2014/15",
      season_id == 27 ~ "2015/16",
      season_id == 37 ~ "2004/05",
      season_id == 38 ~ "2005/06",
      season_id == 39 ~ "2006/07",
      season_id == 40 ~ "2007/08",
      season_id == 41 ~ "2008/09",
      season_id == 42 ~ "2019/20",
      season_id == 90 ~ "2020/21"
      ))

messi_clasico |>
  group_by(season_year) |>
  summarise(
    total = n()
  ) |> arrange(desc(total))
```

```{r}
#This is Messi's shots and goals in Clasico games in La Liga
messi_clasico_sg <- messi_clasico |>
  group_by(season_year) |>
  summarise(
    shots = sum(type.name=="Shot", na.rm = TRUE),
    goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) |> arrange(desc(goals))
```

```{r}
#This is all of his shots and goals by La Liga season
messi_sg <- messi_event |>
  group_by(season_year) |>
  summarise(
    total_shots = sum(type.name=='Shot', na.rm = TRUE),
    total_goals = sum(shot.outcome.name=="Goal", na.rm = TRUE))
```

```{r}
#Here I did a simple bar chart to compare his season stats
messi_sg_wide <- messi_sg |>
  pivot_longer(
    cols = starts_with('total'),
    names_to = "type",
    values_to = "total"
  )

ggplot() + 
  geom_bar(
    data=messi_sg_wide, 
    aes(x=reorder(season_year, total), 
        weight= total,
        fill=type)) + 
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Lionel Messi sees highest number of shots and goals in 2011/12 season",
    y = "Total",
    x = "Season"
  )
```

This ggplot cheat sheet from the guide will prob come in handy later, saving here: <https://raw.githubusercontent.com/rstudio/cheatsheets/master/pngs/data-visualization.png>

```{r}
#Here I worked on a graph from the StatsBomb guide to show all of Messi's completed box passes against Madrid.
library(SBpitch)

messi_passes <- messi_clasico |>
  mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )

messi_event |> allclean()

messi_passes <- messi_passes |>
  filter(type.name=="Pass" & is.na(pass.outcome.name) & player.id==5503) |>
  filter(pass.end_location.x>=102 & pass.end_location.y<=62 & pass.end_location.y >=18)

create_Pitch() +
 geom_segment(data = messi_passes, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Completed Box Passes", subtitle = "La Liga,
vs. Real Madrid") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100)
```

```{r}
#Adding these columns to all dataframes.
messi_clasico <- messi_clasico |>
   mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )

messi_event <- messi_event |>
   mutate(
    pass.end_location.x = map_dbl(pass.end_location, ~ .x[1] %||% NA),
    pass.end_location.y = map_dbl(pass.end_location, ~ .x[2] %||% NA)
  ) |>
  mutate(
    location.x = map_dbl(location, ~ .x[1] %||% NA),
    location.y = map_dbl(location, ~ .x[2] %||% NA)
  )
```

```{r}
#Testing out heat maps!
messi_heatmap <- messi_clasico |> 
  mutate(location.x = ifelse(location.x>120, 120, location.x),
         location.y = ifelse(location.y>80, 80, location.y),
         location.x = ifelse(location.x<0, 0, location.x),
         location.y = ifelse(location.y<0, 0, location.y))

messi_heatmap$xbin <- cut(messi_heatmap$location.x, breaks = seq(from=0, to=120, by = 20),include.lowest=TRUE )
messi_heatmap$ybin <- cut(messi_heatmap$location.y, breaks = seq(from=0, to=80, by = 20),include.lowest=TRUE)

messi_heatmap <- messi_heatmap |>
  filter(type.name=="Shot" | type.name=="Pressure") |>
  group_by(season_year) |>
  mutate(total_sp = n()) |>
  group_by(season_year, xbin, ybin) |>
  summarise(total_sp = max(total_sp),
    bin_sp = n(),
    bin_pct = bin_sp/total_sp,
    location.x = median(location.x),
    location.y = median(location.y)) 

library(grid)

spcolor <- c("#dc2429", "#dc2329", "#df272d", "#df3238", "#e14348", "#e44d51",
"#e35256", "#e76266", "#e9777b", "#ec8589", "#ec898d", "#ef9195",
"#ef9ea1", "#f0a6a9", "#f2abae", "#f4b9bc", "#f8d1d2", "#f9e0e2",
"#f7e1e3", "#f5e2e4", "#d4d5d8", "#d1d3d8", "#cdd2d6", "#c8cdd3", "#c0c7cd",
"#b9c0c8", "#b5bcc3", "#909ba5", "#8f9aa5", "#818c98", "#798590",
"#697785", "#526173", "#435367", "#3a4b60", "#2e4257", "#1d3048",
"#11263e", "#11273e", "#0d233a", "#020c16")


ggplot(data= messi_heatmap, aes(x = location.x, y = location.y, fill = bin_pct, group =bin_pct)) +
 geom_bin2d(binwidth = c(20, 20), position = "identity", alpha = 0.9) + #2
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "white", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "white", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 annotate("point", x = 12 , y = 40, colour = "white", size = 1.05) +
 annotate("point", x = 108 , y = 40, colour = "white", size = 1.05) +
 annotate("path", colour = "white", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 annotate("point", x = 60 , y = 40, colour = "white", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") +
 annotate("path", x=108-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="white") + #4
 scale_y_reverse() + #5
 scale_fill_gradientn(colours = spcolor, trans = "reverse", labels =
scales::percent_format(accuracy = 1), limits = c(0.03, -0.03)) + #6
 labs(title = "TITLE", subtitle = "La Liga, Clasico") + #7
 coord_fixed(ratio = 95/100) + 
  annotation_custom(grob = linesGrob(arrow=arrow(type="open", ends="last",
 length=unit(2.55,"mm")), gp=gpar(col="black", fill=NA, lwd=2.2)),
 xmin=25, xmax = 95, ymin = -83, ymax = -83) + #9
 facet_wrap(~season_year)+ #10
 guides(fill = guide_legend(reverse = TRUE)) + 
  theme(axis.text.x=element_blank(), #remove x axis labels
        axis.ticks.x=element_blank(), #remove x axis ticks
        axis.text.y=element_blank(),  #remove y axis labels
        axis.ticks.y=element_blank()  #remove y axis ticks
        )
```

11/25 Update:

I've tried a few different graphs now to look at overall Messi statistics and then narrow down to performance on the field against Real Madrid. I'm going to test out the same type of chart I did for shots and goals on other things to answer my original question and explore what else can be shown that way. The heat map is a little overwhelming, but I will figure out how to fix it!

```{r}
messi_assists <- messi_clasico |>
  filter(pass.goal_assist == TRUE & player.id==5503) 

create_Pitch() +
 geom_segment(data = messi_assists, aes(x = location.x, y = location.y,
 xend = pass.end_location.x, yend = pass.end_location.y),
 lineend = "round", size = 0.5, colour = "black", arrow =
arrow(length = unit(0.07, "inches"), ends = "last", type = "open")) + 
 labs(title = "Lionel Messi, Assists", subtitle = "La Liga,
vs. Real Madrid") +
scale_y_reverse() + 
coord_fixed(ratio = 105/100)
```

```{r}

shots <- messi_clasico |>
 filter(type.name=="Shot" & (shot.type.name!="Penalty" | is.na(shot.type.name)))

shotmapxgcolors <- c("#192780", "#2a5d9f", "#40a7d0", "#87cdcf", "#e7f8e6", "#f4ef95", "#FDE960", "#FCDC5F",
 "#F5B94D", "#F0983E", "#ED8A37", "#E66424", "#D54F1B", "#DC2608", "#BF0000", "#7F0000", "#5F0000")

ggplot() +
 annotate("rect",xmin = 0, xmax = 120, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 60, ymin = 0, ymax = 80, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 18, xmax = 0, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 102, xmax = 120, ymin = 18, ymax = 62, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = 6, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 114, ymin = 30, ymax = 50, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 120, xmax = 120.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("rect",xmin = 0, xmax = -0.5, ymin =36, ymax = 44, fill = NA, colour = "black", size = 0.6) +
 annotate("segment", x = 60, xend = 60, y = -0.5, yend = 80.5, colour = "black", size = 0.6)+
 annotate("segment", x = 0, xend = 0, y = 0, yend = 80, colour = "black", size = 0.6)+
 annotate("segment", x = 120, xend = 120, y = 0, yend = 80, colour = "black", size = 0.6)+
 theme(rect = element_blank(),
 line = element_blank()) +
 # add penalty spot right
 annotate("point", x = 108 , y = 40, colour = "black", size = 1.05) +
 annotate("path", colour = "black", size = 0.6,
 x=60+10*cos(seq(0,2*pi,length.out=2000)),
 y=40+10*sin(seq(0,2*pi,length.out=2000)))+
 # add centre spot
 annotate("point", x = 60 , y = 40, colour = "black", size = 1.05) +
 annotate("path", x=12+10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40+10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 annotate("path", x=107.84-10*cos(seq(-0.3*pi,0.3*pi,length.out=30)), size = 0.6,
 y=40-10*sin(seq(-0.3*pi,0.3*pi,length.out=30)), col="black") +
 geom_point(data = shots, aes(x = location.x, y = location.y, fill = shot.statsbomb_xg, shape = shot.body_part.name),
 size = 3, alpha = 0.8) +
 labs(title = "Lionel Messi, Shot Map", subtitle = "La Liga, All Clasico Games") + #4
 scale_fill_gradientn(colours = shotmapxgcolors, limit = c(0,0.8), oob=scales::squish, name = "Expected Goals
Value") +
   scale_shape_manual(values = c("Head" = 21, "Right Foot" = 23, "Left Foot" = 24), name ="") + #6
 guides(fill = guide_colourbar(title.position = "top"),
 shape = guide_legend(override.aes = list(size = 4, fill = "black"))) + #7
 coord_flip(xlim = c(85, 125)) +
  theme(axis.text.x=element_blank(),
 axis.title.x = element_blank(),
 axis.title.y = element_blank(),
 plot.caption=element_text(size=5, hjust=0.5, vjust=0.5),
  plot.subtitle = element_text(size = 15, hjust = 0.5),
 axis.text.y=element_blank(), 
 legend.title=element_text(size=10),
 legend.text=element_text(size=10),
 legend.key.size = unit(0.5, "cm"),
 plot.title = element_text(margin = margin(r = 10, b = 10), face="bold",size = 20, colour = "black", hjust = 0.5),
 axis.ticks=element_blank(),
 plot.background = element_rect(fill = "white"),
 strip.text.x = element_text(size=8)) 
```

**12/2 Update**

This week, I worked on two additional graphs. I really like the shot map because I think it answers my initial question of where Messi is shooting from. I also wanted to do more of those plots looking at passing, which is why I went to assists next. I think there's something wrong with one of the arrows because one of them is apparently a corner kick, but it's not shown as such on the graph, so I will take a look at fixing that later. I fear I may just give up on the heat map, or if I want to go back to it, I think I need to narrow down the timeline because it's too much to look at now. I also want to look at the length of the passes possibly. Due to some time constraints and the fact that RStudio kept glitching on me, I will get to those before my final presentation is done. I can also then go back and write a story connecting all of my charts together, taking out my updates and rearranging to move the best charts up top.
